### 2021-04-25 outputs/paper/paper.Rmd
Edit text in paper.Rmd

---
title: "The Impact of Educational Expenses of New Coders"
subtitle: "The Factors Affecting Education Costs of Beginner Programmers to Learn to Code aside from University Tuition"
author: "Bongju Yoo"
thanks: "https://github.com/bonjwow/new-coders"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: ""
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(janitor)
library(stargazer)
library(psych)
# install.packages("here")
library(here)
# install.packages("QuantPsyc")
library(QuantPsyc)
library(car)
library(kableExtra)
# install.packages("forecast")
library(forecast)
```

```{r, include=FALSE}
#### Custom functions ####

### Format number
# Source: https://stackoverflow.com/questions/29465941/format-number-in-r-with-both-comma-thousands-separator-and-specified-decimals

formatNumber <- function(numb, decimal) {
  return(format(round(as.numeric(numb), decimal), nsmall=decimal, big.mark=",")) 
}

### APA style correlation matrix 
# Source: https://stefaneng.github.io/apa_correlation_table/
# install.packages("Hmisc")
library(Hmisc)

apply_if <- function(mat, p, f) {
  # Fill NA with FALSE
  p[is.na(p)] <- FALSE
  mat[p] <- f(mat[p])
  mat
}

apaCorr <- function(mat, corrtype = "pearson") {
  matCorr <- mat
  if (class(matCorr) != "rcorr") {
    matCorr <- rcorr(mat, type = corrtype)
  }
  # Add one star for each p < 0.05, 0.01, 0.001
  stars <- apply_if(round(matCorr$r, 2), matCorr$P < 0.05, function(x) paste0(x, "*"))
  stars <- apply_if(stars, matCorr$P < 0.01, function(x) paste0(x, "*"))
  stars <- apply_if(stars, matCorr$P < 0.001, function(x) paste0(x, "*"))
  # Put - on diagonal and blank on upper diagonal
  stars[upper.tri(stars, diag = T)] <- "-"
  stars[upper.tri(stars, diag = F)] <- ""
  n <- length(stars[1,])
  colnames(stars) <- 1:n
  # Remove _ and convert to title case
  row.names(stars) <- tools::toTitleCase(sapply(row.names(stars), gsub, pattern="_", replacement = " "))
  # Add index number to row names
  row.names(stars) <- paste(paste0(1:n,"."), row.names(stars))
  stars
}
```

# Introduction

With an increasing interest in computer programming and the growth of coding related jobs, more and more people invest their time and money in taking programming courses [@citeHughes]. In this paper, I examine some of the factors that may affect educational expenses of new coders or entry-level programmers aside from university tuition, mainly focusing on the level of education, the size of city, commuting time, income, and learning time. Approximately 20,000 beginner programmers had participated in the survey [@citeSurvey2017].

Firstly, I processed correlation analysis to identify the relationship between each independent variable and dependent variable (education costs) pair and found that all the pairs have a positive linear relationship except for commuting time. Secondly, I used multiple linear regression to determine the statistical significance of the factors. The result of the analysis shows that commuting time is not statistically significant. Interestingly, education background is also not statistically significant associated with educational expenses of beginner programmers. Other variables, such as the size of city, income, and learning time, have a statistical significance on the costs.

This analysis shows that beginner programmers living in small towns and/or low-income households are likely to spend less money on learning to code aside from university tuition than those who live in bigger cities and/or high-income households. It is assumed that there is still a lack of educational resources for new coders living in rural areas and/or low-income households. Free online learning platforms such as freeCodeCamp can be a great resource of education for the coders.

The paper is organized as follows. The Data section describes features of the original survey data and how the data is preprocessed. The Model section explains the multiple linear regression model used to assess the association between each factor and educational costs of beginner programmers, and evaluates the model using its residual standard error, multiple R-squared, and F-statistic. The Results section summarises the results of the regression model and model evaluation processes in the Model section. Lastly, the Discussion section discusses the findings and potential limitations of the paper, and suggests directions for future research related to this data.

# Data

```{r table1, echo=FALSE, warning=FALSE, message=FALSE}
#### Get data ####
dfNewCoders <- 
  readr::read_csv(here::here("inputs/data/clean_new-coders.csv"))

```
## Data Collection
The survey was run by `freeCodeCamp`, a non-profit organization that helps people learn to code through their free online courses and build a network for their alumni [@citeAbout]. The purpose of the survey was to examine how the users learn to code [@cite2017Survey]. The survey was conducted over the internet, and the survey respondents were limited to those persons with less than 5 years learning programming; all the respondents were asked whether they have already been coding for more than 5 years or not before starting the survey [@cite2017Survey]. The survey is composed of 48 questions and takes about five minutes to complete, and the survey results are under the Open Data Common License, which can be freely distributed through the organization’ GitHub repository [@cite2017Survey].^[https://github.com/freeCodeCamp/2017-new-coder-survey]

## Description of Dataset
The original dataset used for this paper is obtained from `freeCodeCamp`’s GitHub page. The format of the dataset is a comma-separated values (CSV) file, which contains 136 columns and 18,175 observations. I selected 8 variables and cleaned the data using the R programming language [@citeR] and the `tidyverse` package [@cityTidyverse] and the `dplyr` package [@cityDplyr]. The selected variables are: Age, CityPopulation, CommuteTime, Gender,  Income, MoneyForLearning, MonthsProgramming, and SchoolDegree. The CityPopulation is the estimated number of city population of the recipient; the question asked was "About how many people live in your city?”, and there are three options to choose: “less than 100,000”, "between 100,000 and 1 million", and "more than 1 million". Since these answers were coded as strings, I converted them into numeric variables using the `recode` function of `dplyr` [@cityDplyr]. The answers for CommuteTime and SchoolDegree were also coded as strings in the original dataset, so I applied the same data cleaning process to the variables as I did with CityPopulation. Also, I omitted observations which have a missing value. Table \@ref(tab:tabViewDataset) is a partial view of the cleaned dataset after the preprocessing.

```{r tabViewDataset, echo=FALSE, warning=FALSE, message=FALSE}
#### A partial view of the cleaned dataset ####
tabCleanedData <-
  dfNewCoders[1:10,] %>%
  mutate_all(linebreak) %>%
  kable(caption = "A partial view of the cleaned dataset",
      booktabs = T,
      escape = F,
      "latex",
      col.names = linebreak(c("Gender", 
                              "Age",
                              "City\nPopulation",
                              "Commute\nTime",
                              "Income",
                              "Months\nProgramming",
                              "School\nDegree",
                              "Money For\nLearning"), align = "c")) %>%
  kable_classic(full_width = F)

tabCleanedData

```

Table 2 displays descriptive statistics for the cleaned data after the preprocessing. The total number of observations is 7,022 and the data type for all variables are numeric variables. The median age of the respondents is 30 years and the median income of them is about \$43,000 in U.S. dollars. The average of the respondents has spent about 2 years and about \$1,000 in U.S. dollars learning to code.

```{r tabDescrStat, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
#### Descriptive statistics ####
stargazer::stargazer(data.frame(dfNewCoders), 
                     type = "latex", 
                     title = "Descriptive statistics for the cleaned dataset", 
                     header = FALSE,
                     single.row = TRUE)

```

## Correlation Analysis

I ran a correlation matrix to check correlation coefficients between the variables. To create the correlation matrix, I used a custom function written by Stefan Eng [@citeEng]. His function contains the `Hmisc` package’s `rcorr` function to calculate the correlations with the p-values and display starts based on the significance level of each cell when the value is lower than a specific level [@citeHmisc]. Along with his custom to compute the correlation matrix, I used the `kable` package to print out the result in a table format [@citeKable]. As can be seen in Table \@ref(tab:tabCorrAnalysis), the result of the correlation matrix shows that all independent variables have a positive significant relationship with the dependent variable, MoneyForLearning, except for CommuteTime. This can be interpreted that a commuting time to work does not affect educational expenditure of new coders aside from their post-secondary education cost.

```{r tabCorrAnalysis, echo=FALSE, warning=FALSE, message=FALSE}
#### Correlation analysis ####

### Print correlation coefficient
tblCorr <-
  dfNewCoders %>%
  as.matrix() %>%
  apaCorr() %>%
  data.frame() %>%
  kable(caption = "Correlation matrix for variables",
        booktabs = TRUE,
        col.names = c("1", "2", "3", "4", "5", "6", "7", "8")) %>%
  footnote(general = "* p < 0.05, ** p < 0.01, *** p < 0.001",
             general_title = "")

tblCorr

# No correlation coefficient > 0.9 
# No multicollinearity
```

# Model
```{r, include=FALSE}
#### TODO's ####
# Write out what each variable is
# Weaknesses and next steps
```

## Model Formulae

In the earlier section, I ran a correlation matrix to identify the correlation of each variable and the significance of the relationship between the dependent variable (MoneyForLearning) and other dependents variables. But, I was not able to estimate how much the independent variables impact the dependent variable from the correlation analysis. In this section, I will examine effects of the dependent variables on the dependent variable, focusing on one of the dependent variables: Income. The two main questions that I set out to answer in this section are: (1) How do beginner programmers' income affect the cost of their education to learn code? and (2) Will the effects change according to the population of the city where they live in.

To obtain the answer for the first question, I used a simple linear regression model (Equation \@ref(eq:linRegModel1)) and multiple linear regression models (Equation \@ref(eq:linRegModel2) and \@ref(eq:linRegModel3)). All these three models use MoneyForLearning (the cost of their education to learn code) as a dependent variable. Model 1 contains only one independent variable, which is Income (see Equation \@ref(eq:linRegModel1)); otherwise, Model 3 uses all the independent options except for CommuteTime: Gender, Age, CityPopulation, Income, MonthsProgramming, and SchoolDegree (see Equation \@ref(eq:linRegModel3)). As mentioned in the Correlation Analysis section, the CommuteTime (a commuting time) variable does not significantly affect the response variable, so it is excluded from the regression models. And, Model 2 uses the same independent variables, excluding Gender and Age. For the second question about the effects of the population of the city where the programmers reside, I used four simple linear regression models (see Equation \@ref(eq:linRegModel2)). Like Model 1, all these models use MoneyForLearning as a dependent variable and Income as an independent variable (see Equation \@ref(eq:linRegModel1)). But, the observations were grouped by the population size of the programmers. Model 4 contains any size of population; and Model 5,6, and 7 are the small (less than 100,000), medium (between 100,000 and 1 million), and large (more than 1 million) size population.

\begin{equation}
(\#eq:linRegModel1)
Y \sim \beta_0 + \beta_1X_1 + \epsilon
\end{equation}

\begin{equation}
(\#eq:linRegModel2)
Y \sim \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 + \beta_4X_4 + \epsilon
\end{equation}

\begin{equation}
(\#eq:linRegModel3)
Y \sim \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 + \beta_4X_4 + \beta_5X_5 + \beta_6X_6 + \epsilon
\end{equation}

## Data Preprocessing

As explained in the Data section, categorical variables were converted to continuous variables for the linear regression models. For example, the SchoolDegree variable (the level of education) was a categorical variable with character values. So, I converted the variable into a variable with numeric values, using the `dplyr` package. A complete list of survey questions and response options for each question can be found in the Appendix. Since the MoneyForLearning variable (the level of education) was not only categorical but also ordered and ranked in the original dataset, I converted the variable into a variable with numeric values. Also, I merged some of the response options of the variable. For instance, it is somewhat hard to say that a high school diploma or equivalent (GED) is higher than trade, technical, or vocational training, and a professional degree, such as MBA, MD, or JD, is higher than a non-professional master's degree. So, I merged the options, using the `recode` function of the `dplyr` package [@cityDplyr].

## Model Validation

I will use Mean Square Error (MSE) and Root Mean Square Error (RMSE) to evaluate the accuracy of the model. As the name implies, MSE is defined as the mean of the square of the residuals obtained from a regression model, and RMSE is the square root of MSE [@citeHolmes]. I will use the residuals which are the results of the `lm` function to calculate MSE and RMSE of the models. Firstly, I will calculate the square of the residuals, and then computed the mean of the squared value for MSE (see Equation \@ref(eq:mse)). Secondly, for RMSE, I will use the `sqrt` function to calculate the square root of MSE (see Equation \@ref(eq:rmse)). Once I obtain the results of MSE and RMSE, I will compare them to see how accurately the models predict the response.

\begin{equation}
(\#eq:mse)
MSE = \frac{1}{N}\sum_{i=1}^{N} (residual_{i})^2
\end{equation}

\begin{equation}
(\#eq:rmse)
RMSE = \sqrt{\frac{1}{N}\sum_{i=1}^{N} (residual_{i})^2}
\end{equation}

Since two multiple linear regression models (Model 2 and 3) are used in the first group of regression models, I will check if multicollinearity exists in these regression models, using variance inflation factor (VIF) and tolerance. I will use the `car` package to calculate the values of VIF and tolerance for each model [@citeCar]. Also, I will identify how much each of the independent variables affects the dependent variable with the standardized regression coefficients (beta). the `QuantPsyc` package will be used to compute the beta value [@citeQuanPsy].

# Results

## Results of Regression Model

All the regression models in the first group of the models had p-values of F-statistic lower than a significance level of 0.05 (see Table 4). And, each of the independent variables also had a p-value of less than 0.05 except for the SchoolDegree variable in Model 3. This can be interpreted that all the independent variables excluding the SchoolDegree variable have an impact on the response variable (MoneyForLearning), and the level of education (SchoolDegree) is not statistically significant on the response variable with holding all other variables. However, the SchoolDegree variable becomes statistically significant if the Gender and Age variables are excluded from the regression model.

```{r tabLinRegModel1, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
#### Linear regression models (Group 1) ####

### Run linear regression models
multiReg1 <- lm(formula = 
               MoneyForLearning ~ 
               Income,
               data = dfNewCoders)

multiReg2 <- lm(formula = 
               MoneyForLearning ~ 
               # Gender +
               # Age +
               CityPopulation +
               # CommuteTime +
               Income +
               MonthsProgramming +
               SchoolDegree, 
               data = dfNewCoders)

multiReg3 <- lm(formula = 
               MoneyForLearning ~ 
               Gender +
               Age +
               CityPopulation +
               # CommuteTime +
               Income +
               MonthsProgramming +
               SchoolDegree, 
               data = dfNewCoders)

# summary(multiReg3)
stargazer::stargazer(multiReg1, multiReg2, multiReg3, 
          title = "Predictions of money for learning", 
          header = FALSE,
          type = "latex",
          column.labels = c("Model (1)", "Model (2)", "Model (3)"),
          model.numbers = FALSE)
```

The second group of regression models to examine the effects of the population of the city where the programmers reside shows that all the models have statistically significant p-values of F-statistic except for Model 5, the model for the small city with less than 100,000 residents (see Table 5). In other words, beginner programmers' income affects the cost of their education to learn code; however, their income does not affect the cost when the model contains only observations from the small city.

```{r tabLinRegModel2, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
#### Linear regression models (Group 2) ####
dfSmCity <-
  dfNewCoders %>%
  filter(CityPopulation == 0)

dfMdCity <-
  dfNewCoders %>%
  filter(CityPopulation == 1)

dfLgCity <-
  dfNewCoders %>%
  filter(CityPopulation == 2)

### Run linear regression models
linReg1 <- lm(formula = 
               MoneyForLearning ~ 
               Income,
               data = dfNewCoders)

linReg2 <- lm(formula = 
               MoneyForLearning ~ 
               Income,
               data = dfSmCity)

linReg3 <- lm(formula = 
               MoneyForLearning ~ 
               Income,
               data = dfMdCity)

linReg4 <- lm(formula = 
               MoneyForLearning ~ 
               Income,
               data = dfLgCity)

stargazer::stargazer(linReg1, linReg2, linReg3, linReg4,
          title = "Predictions of money for learning by city population", 
          header = FALSE,
          type = "latex",
          column.labels = c("Model (4): Any", "Model (5): Small", "Model (6): Medium", "Model (7): Large"),
          model.numbers = FALSE,
          font.size = "tiny")

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#### RMSE with custom function ####

# Source: https://medium.com/dev-genius/metrics-for-evaluating-linear-regression-models-36df305510d9
# Source: https://stackoverflow.com/questions/26237688/rmse-root-mean-square-deviation-calculation-in-r

### Define functions for R-squared, MSE, and RMSE
RSQ <- function(model) {
  result <- summary(model)$r.squared
  return(formatNumber(result, 3))
}

MSE <- function(error) {
  result <- mean(error^2)
  return(formatNumber(result, 3))  
}

RMSE <- function(error) {
  result <- sqrt(mean(error^2))
  return(formatNumber(result, 3)) 
}

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

### Calculate R-squared, MSE, and RMSE for each model 
vecModel <-
  c("Model (1)", "Model (2)", "Model (3)")

vecRSQ <- 
  c(RSQ(multiReg1), RSQ(multiReg2), RSQ(multiReg3))

vecMSE <-
  c(MSE(multiReg1$residuals), MSE(multiReg2$residuals), MSE(multiReg3$residuals))

vecRMSE <-
  c(RMSE(multiReg1$residuals), RMSE(multiReg2$residuals), RMSE(multiReg3$residuals))

### Display the result in a table
dfModelValid <-
  data.frame(vecModel, vecRSQ, vecMSE, vecRMSE) %>%
  kable(caption = "Results of model validation metrics for model group 1",
        booktabs = TRUE,
        col.names = c("Model", "R-squared", "MSE", "RMSE")) %>%
  column_spec(column = 1, width = "4cm") %>%
  column_spec(column = 2:4, width = "3cm")

dfModelValid

# If R-squared is closer to 0 = Good!
# Model3 has the lowest RMSE among the 3 models => Model3 is the best!
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

### Calculate R-squared, MSE, and RMSE for each model
vecModel_city <-
  c("Model (4)", "Model (5)", "Model (6)", "Model (7)")

vecRSQ_city <- 
  c(RSQ(linReg1), RSQ(linReg2), RSQ(linReg3), RSQ(linReg4))

vecMSE_city <-
  c(MSE(linReg1$residuals), MSE(linReg2$residuals), MSE(linReg3$residuals),  MSE(linReg4$residuals))

vecRMSE_city <-
  c(RMSE(linReg1$residuals), RMSE(linReg2$residuals), RMSE(linReg3$residuals), RMSE(linReg4$residuals))

### Display the result in a table
dfModelValidCity <-
  data.frame(vecModel_city, vecRSQ_city, vecMSE_city, vecRMSE_city) %>%
  kable(caption = "Results of model validation metrics for model group 2",
        booktabs = TRUE,
        col.names = c("Model", "R-squared", "MSE", "RMSE")) %>%
  column_spec(column = 1, width = "4cm") %>%
  column_spec(column = 2:4, width = "3cm")

dfModelValidCity

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#### Multicollinearity diagnostics ####
# Check multicollinearity with Variance Inflation Factor (VIF) & Tolerance

multicolDiagno <- function(mode, title) {
  betaModel2 <- QuantPsyc::lm.beta(mode)
  vifModel2 <- car::vif(mode)
  tolModel2 <- 1/car::vif(mode)

  dfMulticol <-
    betaModel2 %>%
    rbind(formatNumber(unname(betaModel2), 3)) %>%
    rbind(formatNumber(unname(vifModel2), 3)) %>%
    rbind(formatNumber(unname(unname(tolModel2)), 3)) %>%
    cbind(c("B", "Beta", "VIF", "Tolerance"))
    # dplyr::relocate(Metric, .before = CityPopulation) %>%
    # dplyr::select(Metric, everything()) %>%
    # kable()

  ### Relocate 'Metric' column: alternative to 'dplyr::relocate()'
  dfMulticol <- dfMulticol[-1,]
  dfMulticol <- dfMulticol[,c(ncol(dfMulticol),1:ncol(dfMulticol)-1)]

  kable(dfMulticol, 
        caption = title,
        booktabs = TRUE)
}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#### Multicollinearity diagnostics for Model (2) ####
multicolDiagno(multiReg2, "Results of multicollinearity diagnostic metrics for Model (2)")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#### Multicollinearity diagnostics for Model (3) ####
multicolDiagno(multiReg3, "Results of multicollinearity diagnostic metrics for Model (3)")
```

# Discussion
```{r, include=FALSE}
#### TODO's ####
# Write at least 2.5 pages for the Discussion section
# Weaknesses and next steps
```

\newpage

# Appendix {-}
## Survey questions {-}

Q. [Age] How old are you?

* Type: integer

Q. [CityPopulation] About how many people live in your city?

* Type: string

* Options:
  1. less than 100,000
  2. between 100,000 and 1 million
  3. more than 1 million

Q. [CommuteTime] About how many minutes does it take you to get to work each day?

* Type: string

* Options:
  1. I work from home
  2. Less than 15 minutes
  3. 15 to 29 minutes
  4. 30 to 44 minutes
  5. 45 to 60 minutes
  6. More than 60 minutes

Q. [Gender] What's your gender?

* Type: string

* Options:
  1. male
  2. female

Q. [Income] About how much money did you make last year, in US dollars?

* Type: integer

Q. [MoneyForLearning] Aside from university tuition, about how much money have you spent on learning to code so far, in US dollars?

* Type: integer

Q. [MonthsProgramming] About how many months have you been programming for?

* Type: integer

Q. [SchoolDegree] What's the highest degree or level of school you have completed?

* Type: string

* Options:
  1. no high school (secondary school)
  2. some high school
  3. high school diploma or equivalent (GED)
  4. trade, technical, or vocational training
  5. some college credit, no degree
  6. associate's degree
  7. bachelor's degree
  8. master's degree (non-professional)
  9. professional degree (MBA, MD, JD, etc.)
  10. Ph.D.

\newpage

# References